A Data-Driven Framework for Neural Field Modeling

### 2011-01-18 ltx/first_order_framework/response_to_reviews.tex

    \end{enumerate}

    \subsection{Additional Comments}

    \begin{enumerate}
        \item The discrepancy of the results in Experiment II where the weights of the basis functions are not in agreement with the actual parameters while the reconstructed kernels seem satisfying.

		\emph{We have added to the results section describing experiment II: `These results indicate that the solution to the system may not be unique in non-ideal, practical situations. Nevertheless, the net shape of the connectivity kernel can be determined reasonably accurately.'}

@@

    \end{enumerate}

    \section{Reviewer 2}

\begin{enumerate}
    \item The handling of the (to borrow a term from numerical weather forecasting) `sub-grid scale physics' is inadequate for nervous systems (see, e.g., equation 8 (now equation 12) [The nonlinear IntegroDifference Equation]). I do very much appreciate the spectral estimation of spatial and temporal scales. But the model errors are lumped into an iid model error term that is not realistic. For instance, the flow of potassium in the extracellular spaces has a known dynamic that will affect network excitability, generating important sub-grid dynamics that is anything but iid. See, for instance, the example in Ullah et al PLoS Comp Biol 2010 where without such dynamics, state tracking and parameter estimation for a neural system fails. A beautiful overview of sub-grid effects in data assimilation is found in Kalnay's 2003 book. I would very much value the authors comments on these issues. I guess were I to respond to this question, that I might consider extending the model error term to an autoregressive model. But I will let the authors answer this question.

	\emph{A more sophisticated model for the error will provide a more realistic description of the cortical dynamics. However, this will increase the number of parameters making the estimation problem more difficult. Furthermore, the authors believe that is a benefit staying as close as possible to the classic WC equations, so future results can be compared to theoretical studies. A preferable alternative to an AR error term is to allow for time-varying parameters to account for changes in excitability levels. In future work, we also hope estimate the disturbance properties (spatial covariance and temporal variance). To address this point in this paper, we have added an extra paragraph to `Neural Field Model' section of the Methods, where we discuss the model mismatch and some basic modeling assumptions, such as stationarity of the parameters. We have also added to the Estimation section, stating that we are assuming the parameters are stationary of the estimation period. In addition, we have added to the discussion (Extensions section) with possible ways to relax these assumptions.}

### 2011-01-18 ltx/first_order_framework/response_to_reviews.tex

\item I understand that the Mexican Hat bases are in this paper often a composite of $\theta_0$, $\theta_1$, and $\theta_2$, representing short range excitation, short range inhibition, and long range excitation. But many will be confused by this subtlety, and at times in the text the Mexican Hat basis is described as if it were a single function. I would show how the theta's sum to create the general spatial basis shape of the hat in figure 1, and go through the paper in order to clarify this issue.

\emph{This point is clarified by adding the connectivity kernel components superimposed by the resulting Mexican Hat basis in Figure 1. of the manuscript. It has also been clarified throughout the paper.}
 % 
 % \begin{figure*}[!ht]
 % \begin{center}
 % \includegraphics{./Graph/pdf/fig1.pdf} 
 % \end{center}
 % \caption{}
 % \label{fig:Figure1}
 % \end{figure*}

\item A few comments on how to adapt the smoother for real time implementation would be helpful. I presume that one would perform the smoothing during a training phase. But one would also wish to adapt in real time as observation or control scenarios progressed. A few words of perspective would be invaluable to help others translate the author's wisdom on this method to other applications.

\emph{A few comments have been added to the `Extensions to the Framework' section of the discussion:\\\\ 'The estimation framework that we have presented is inherently a batch algorithm designed to operate after data has been collected. A further extension of the framework is to allow for the algorithm to learn online. A filter that is designed to be implemented in an online setting can be used in place of the smoother along with an online parameter estimator to update the parameter after a new observation arrived. A real-time implementation can be achieved using the UKF (instead of the URTSS) and an online recursive LS estimator Ljung (1999).'}

### 2011-01-18 ltx/first_order_framework/LaTeX_Manuscript.tex
edits to response in AR error bit
edit the discretization appendix

% If $0 < T_s < \tau$, where the synaptic time constant $\tau = 1/\zeta$, the IDE field model is stable. To show this, the model is written as the series
% \begin{align}
% 	v_1(\mathbf{r}) &= \xi v_0(\mathbf{r}) + T_s g_0(\mathbf{r}) \\
% 	v_2(\mathbf{r})	&= \xi^2 v_0(\mathbf{r}) + T_s \left(\xi g_0(\mathbf{r}) + g_1(\mathbf{r}) \right) \\
% 	\vdots \, \, \, \, \, &= \, \, \, \, \, \vdots \nonumber \\
% 	v_{t+1}(\mathbf r) &= \xi^{t+1} v_0\left(\mathbf{r} \right) + T_s \sum_{i=0}^t \xi^i g_{t-i}(\mathbf{r}), \label{eq:fieldmodelseries}
% \end{align}
% where
% \begin{equation}
% 	g_{t}(\mathbf{r}) = \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}.
% \end{equation}
% Since  $0 \leq f(v_t(\mathbf r)) \leq 1$ and the kernel is semi-compactly supported, $g_t(\mathbf r)$ is always bounded . To assess the stability of the system we assume that $g_t(\mathbf{r})$ is maximally saturated over the entire field for all time, which we denote as $g_{max}$. Substituting this into equation~\ref{eq:fieldmodelseries} yields
% \begin{align}\label{vrecursiveinlimit}
% 	v_{t+1}(\mathbf r) &= \xi^{t+1}v_0(\mathbf r)+ g_{max} T_s \sum_{i=0}^t \xi^i \nonumber \\
% 	&=\xi^{t+1}v_0(\mathbf r) + \frac{ g_{max} T_s(1-\xi^{t+1})}{1-\xi}.
% \end{align}
% To study the asymptotic behaviour of the system we consider the limit as $t$ approaches infinity. Since $0 < \xi <1$  we use a property of the geometric series to simplify equation \eqref{vrecursiveinlimit} to
% \begin{equation}\label{simplifiedvrecursiveinlimit}
% 	v_{t+1}(\mathbf r)=\xi^{t+1}v_0(\mathbf r)+\frac{g_{max} T_s}{1-\xi}.
% \end{equation}
% Substituting $\xi=1-\zeta T_s$ into equation \eqref{simplifiedvrecursiveinlimit} gives
% \begin{align}
% 	v_{t+1}(\mathbf r) &= (1-\zeta T_s)^{t+1}v_0(\mathbf r)+\frac{g_{max}}{\zeta} \\
% 	&= \left(1-\frac{T_s}{\tau}\right)^{t+1}v_0(\mathbf r) + \tau g_{max}
% \end{align}
% From this equation it can be seen that if $0 < T_s < \tau$, the system is stable. The sampling period was set to $T_s = 0.001 s$ and the time constant was set to $\tau = 0.01 s^{-1}$. Therefore, the discretization preserved the stability properties of the system. 
Note, the discretization using the first Euler method has the potential to affect the system's behaviour. If the sampling period is too large, the difference between the continuous and discrete-time systems can be substantial. The sampling period of 1~ms was used in this paper, which was sufficiently small to yield stable discrete-time system and good estimation results. An analysis of the error introduced by Euler discretization method can be found in \citet{Butcher2008}. An alternative to the Euler's method is the local linearization method, which may provide improved results~\citep{Ozaki1994}.

>>>>

% If $0 < T_s < \tau$, where the synaptic time constant $\tau = 1/\zeta$, the IDE field model is stable. To show this, the model is written as the series
% \begin{align}
% 	v_1(\mathbf{r}) &= \xi v_0(\mathbf{r}) + T_s g_0(\mathbf{r}) \\
% 	v_2(\mathbf{r})	&= \xi^2 v_0(\mathbf{r}) + T_s \left(\xi g_0(\mathbf{r}) + g_1(\mathbf{r}) \right) \\
% 	\vdots \, \, \, \, \, &= \, \, \, \, \, \vdots \nonumber \\
% 	v_{t+1}(\mathbf r) &= \xi^{t+1} v_0\left(\mathbf{r} \right) + T_s \sum_{i=0}^t \xi^i g_{t-i}(\mathbf{r}), \label{eq:fieldmodelseries}
% \end{align}
% where
% \begin{equation}
% 	g_{t}(\mathbf{r}) = \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}.
% \end{equation}
% Since  $0 \leq f(v_t(\mathbf r)) \leq 1$ and the kernel is semi-compactly supported, $g_t(\mathbf r)$ is always bounded . To assess the stability of the system we assume that $g_t(\mathbf{r})$ is maximally saturated over the entire field for all time, which we denote as $g_{max}$. Substituting this into equation~\ref{eq:fieldmodelseries} yields
% \begin{align}\label{vrecursiveinlimit}
% 	v_{t+1}(\mathbf r) &= \xi^{t+1}v_0(\mathbf r)+ g_{max} T_s \sum_{i=0}^t \xi^i \nonumber \\
% 	&=\xi^{t+1}v_0(\mathbf r) + \frac{ g_{max} T_s(1-\xi^{t+1})}{1-\xi}.
% \end{align}
% To study the asymptotic behaviour of the system we consider the limit as $t$ approaches infinity. Since $0 < \xi <1$  we use a property of the geometric series to simplify equation \eqref{vrecursiveinlimit} to
% \begin{equation}\label{simplifiedvrecursiveinlimit}
% 	v_{t+1}(\mathbf r)=\xi^{t+1}v_0(\mathbf r)+\frac{g_{max} T_s}{1-\xi}.
% \end{equation}
% Substituting $\xi=1-\zeta T_s$ into equation \eqref{simplifiedvrecursiveinlimit} gives
% \begin{align}
% 	v_{t+1}(\mathbf r) &= (1-\zeta T_s)^{t+1}v_0(\mathbf r)+\frac{g_{max}}{\zeta} \\
% 	&= \left(1-\frac{T_s}{\tau}\right)^{t+1}v_0(\mathbf r) + \tau g_{max}
% \end{align}
% From this equation it can be seen that if $0 < T_s < \tau$, the system is stable. The sampling period was set to $T_s = 0.001 s$ and the time constant was set to $\tau = 0.01 s^{-1}$. Therefore, the discretization preserved the stability properties of the system.

Note, the discretization using the first order Euler method has the potential to affect the system's behaviour, where the difference between the continuous and discrete-time systems can be substantial. The Euler method was used for its simplicity. The sampling period of 1~ms was used in this paper, which yielded a stable discrete-time system and good estimation results. An analysis of the error introduced by Euler discretization method can be found in~\citet{Butcher2008}. An alternative to Euler's method is the local linearization method~\citep{Ozaki1994}.
